Reproduction frequency and offspring survival decline in elephant seals past prime age

### 2024-10-17 paper/paperv3.qmd
updating model sensitivity to threshold analyses with h4 phenology

  effect_size <- function(m) {
    probs <- tibble(age = c(15, 19),
                    age_cat = c("Old", "Old")) %>% 
      mutate(age10 = (age - a) / 10) %>% 
      predict(m, re.form = NA, newdata = ., type = "response")
    odds <- probs / (1 - probs)
    odds[2] / odds[1]
  }

  list(
    h1 = summary(breed_age_mod)$coefficients,
    h2_surv = summary(pup_surv_mod)$coefficients,
    h2_recr = summary(pup_recr_mod)$coefficients,
    h3 = summary(sexrat_age_mod)$coefficients,
    h1_aic = AIC(breed_age_mod),
    h2_surv_aic = AIC(pup_surv_mod),
    h2_recr_aic = AIC(pup_recr_mod),
    h3_aic = AIC(sexrat_age_mod),
    h1_effect = effect_size(breed_age_mod),
    h2_surv_effect = effect_size(pup_surv_mod),
    h2_recr_effect = effect_size(pup_recr_mod),
    h3_effect = effect_size(sexrat_age_mod)
  )
}

@@

# Are hypotheses supported by the model? note to future readers...this part could have been done way easier, and you will have to trust our process that this was the path of least resistence for us at the point of final revisions!
extract_p <- function(h, tails = 1) {
  if (tails == 1) {
    map_dbl(mod_sens, \(m) pnorm(m[[h]]["age10:age_catOld", "z value"]))
  } else if (tails == 2) {
    map_dbl(mod_sens, \(m) 2 * pnorm(m[[h]]["age10:age_catOld", "z value"],
                                     lower.tail = (m[[h]]["age10:age_catOld", "z value"] < 0)
                                     ))
  } else {
    stop("tails must be 1 or 2")
  }
}
extract_aic <- function(h) {
  map_dbl(mod_sens, paste(h, "aic", sep = "_"))
}
p_tbl <- tibble(age_cutoff = cutoff, 
                "H1 (breeding)" = extract_p("h1", tails = 2),
                "H2a (survival)" = extract_p("h2_surv", tails = 2),
                "H2b (recruitment)" = extract_p("h2_recr", tails = 2),
                "H3 (sex ratio)" = extract_p("h3", tails = 2))
aic_weight <- function(x) {
  rellik <- exp(-0.5 * (x - min(x)))
  rellik / sum(rellik)
}
aic_tbl <- tibble(age_cutoff = cutoff,
                  "H1 (breeding)" = extract_aic("h1"),
                  "H2a (survival)" = extract_aic("h2_surv"),
                  "H2b (recruitment)" = extract_aic("h2_recr"),
                  "H3 (sex ratio)" = extract_aic("h3")) %>% 
  mutate(across(-age_cutoff, aic_weight))

extract_coef <- function(h) {
  map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Estimate"])
}
coef_tbl <- tibble(age_cutoff = cutoff, 
                "H1 (breeding)" = extract_coef("h1"),
                "H2a (survival)" = extract_coef("h2_surv"),
                "H2b (recruitment)" = extract_coef("h2_recr"),
                "H3 (sex ratio)" = extract_coef("h3"))

>>>>

  # H4
molt_data <- sealdat %>% 
  filter(year >= 2011, 
         moltdur >= 7, 
         age <= 20) %>% 
  mutate(BA = firstobsbreeddoy, 
         BD = firstobsbreeddoy + breeddur, 
         MA = lastobsmoltdoy - moltdur, 
         MD = lastobsmoltdoy) %>% 
  drop_na(age, BA, BD, MA, MD) %>% 
  transmute(animalID, 
            year_fct, 
            age10 = (age - age_senesce) / 10, 
            age_cat = factor(age >= age_senesce, 
                          labels = c("Young", "Old")),
            moltdur, 
            breeddur) %>% 
  pivot_longer(moltdur:breeddur, 
               names_to = "phase", 
               values_to = "dur_days") %>% 
  filter(phase == "moltdur")

molt_mod <- glmer(dur_days ~ age10 : age_cat + (1 | year_fct) + (1 | animalID), 
        molt_data, 
        family = Gamma(link = "log"))

breed_data <- sealdat %>% 
  filter(year >= 2011, 
         moltdur >= 7, 
         age <= 20) %>% 
  mutate(BA = firstobsbreeddoy, 
         BD = firstobsbreeddoy + breeddur, 
         MA = lastobsmoltdoy - moltdur, 
         MD = lastobsmoltdoy) %>% 
  drop_na(age, BA, BD, MA, MD) %>% 
  transmute(animalID, 
            year_fct, 
            age10 = (age - age_senesce) / 10, 
            age_cat = factor(age >= age_senesce, 
                          labels = c("Young", "Old")),
            moltdur, 
            breeddur) %>% 
  pivot_longer(moltdur:breeddur, 
               names_to = "phase", 
               values_to = "dur_days") %>% 
  filter(phase == "breeddur")

breed_mod <- glmer(dur_days ~ age10 : age_cat + (1 | year_fct) + (1 | animalID), 
        data = breed_data, 
        family = Gamma(link = "log"), 
        control = glmerControl(optimizer = "bobyqa")
)

#Making the tables
#Not sure about this function
  # effect_size <- function(m) {
  #   probs <- tibble(age = c(15, 19),
  #                   age_cat = c("Old", "Old")) %>% 
  #     mutate(age10 = (age - a) / 10) %>% 
  #     predict(m, re.form = NA, newdata = ., type = "response")
  #   odds <- probs / (1 - probs)
  #   odds[2] / odds[1]
  # }

  list(
    h1 = summary(breed_age_mod)$coefficients,
    h2_surv = summary(pup_surv_mod)$coefficients,
    h2_recr = summary(pup_recr_mod)$coefficients,
    h3 = summary(sexrat_age_mod)$coefficients,
    h4_molt = summary(molt_mod)$coefficients, 
    h4_breed = summary(breed_mod)$coefficients,
    h1_aic = AIC(breed_age_mod),
    h2_surv_aic = AIC(pup_surv_mod),
    h2_recr_aic = AIC(pup_recr_mod),
    h3_aic = AIC(sexrat_age_mod),
    h4_molt_aic = AIC(molt_mod), 
    h4_breed_aic = AIC(breed_mod)
    # h1_effect = effect_size(breed_age_mod),
    # h2_surv_effect = effect_size(pup_surv_mod),
    # h2_recr_effect = effect_size(pup_recr_mod),
    # h3_effect = effect_size(sexrat_age_mod)
  )
}

@@

```{r}
# Are hypotheses supported by the model? note to future readers...this part could have been done way easier, and you will have to trust our process that this was the path of least resistence for us at the point of final revisions!
# extract_p <- function(h, tails = 1) {
#   if (tails == 1) {
#     map_dbl(mod_sens, \(m) pnorm(m[[h]]["age10:age_catOld", "z value"]))
#   } else if (tails == 2) {
#     map_dbl(mod_sens, \(m) 2 * pnorm(m[[h]]["age10:age_catOld", "z value"],
#                                      lower.tail = (m[[h]]["age10:age_catOld", "z value"] < 0)
#                                      ))
#   } else {
#     stop("tails must be 1 or 2")
#   }
# }
extract_p <- function(h) {
  map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Pr(>|z|)"])
}
extract_aic <- function(h) {
  map_dbl(mod_sens, paste(h, "aic", sep = "_"))
}
aic_weight <- function(x) {
  rellik <- exp(-0.5 * (x - min(x)))
  rellik / sum(rellik)
}
extract_coef <- function(h) {
  map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Estimate"])
}
extract_se <- function(h) {
  map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Std. Error"])
}

#make the tables
p_tbl <- tibble(age_cutoff = cutoff, 
                "H1 (breeding)" = extract_p("h1"),
                "H2a (survival)" = extract_p("h2_surv"),
                "H2b (recruitment)" = extract_p("h2_recr"),
                "H3 (sex ratio)" = extract_p("h3"), 
                "H4a (molt duration)" = extract_p("h4_molt"), 
                "H4b (breed duration)" = extract_p("h4_breed"))

aic_tbl <- tibble(age_cutoff = cutoff,
                  "H1 (breeding)" = extract_aic("h1"),
                  "H2a (survival)" = extract_aic("h2_surv"),
                  "H2b (recruitment)" = extract_aic("h2_recr"),
                  "H3 (sex ratio)" = extract_aic("h3"), 
                  "H4a (molt duration)" = extract_aic("h4_molt"), 
                  "H4b (breed duration)" = extract_aic("h4_breed")) %>% 
  mutate(across(-age_cutoff, aic_weight))

coef_tbl <- tibble(age_cutoff = cutoff, 
                "H1 (breeding)" = extract_coef("h1"),
                "H2a (survival)" = extract_coef("h2_surv"),
                "H2b (recruitment)" = extract_coef("h2_recr"),
                "H3 (sex ratio)" = extract_coef("h3"), 
                "H4a (molt duration)" = extract_coef("h4_molt"), 
                "H4b (breed duration)" = extract_coef("h4_breed"))

se_tbl <- tibble(age_cutoff = cutoff, 
                "H1 (breeding)" = extract_se("h1"),
                "H2a (survival)" = extract_se("h2_surv"),
                "H2b (recruitment)" = extract_se("h2_recr"),
                "H3 (sex ratio)" = extract_se("h3"), 
                "H4a (molt duration)" = extract_se("h4_molt"), 
                "H4b (breed duration)" = extract_se("h4_breed"))
